---
title: "Assignment 1 - Economic History"
output:
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Packages and Environment

```{r, results='hide'}
library(openxlsx)
library(tidyr)
library(dplyr)
library(lfe)
library(fixest)
library(jsonlite)

rm(list=ls())
setwd("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History")

```

### 2.  Download the Excel files

```{r}
car = read.xlsx("StateNewCarRegistrations.xlsx")
stock = read.xlsx("stock_income.xlsx")

head(car)
head(stock)
```

### 3. Create a single dataset.

```{r}
car_long <- pivot_longer(car, #we start by pivoting the car dataset
                               cols = -c(month, year), 
                               names_to = "state", 
                               values_to = "car_sales")
head(car_long)



dta3 <- merge(stock, car_long, by = "state", all.x = TRUE)%>%
  filter(!is.na(year)) %>% #We merge both datasets by state, filter for NAs (and remove the "total" observations)
  group_by(state)%>% #We group observations by state
  mutate(month = as.numeric(month)) %>% #We turn the class of month from character to numeric
  arrange(year, month, .by_group = TRUE) #This allows us to arrange observations by year and month within state groups

head(dta3) #The dataset seems to corresponds to the exemple given.
```


### 4. Create a post-crash dummy

```{r}
dta4 <- dta3 %>%
  mutate (crash = ifelse(year < 1929 | (year == 1929 & month <= 10), 0, 1))
```


### 5.  In an ideal world, what should $x_s$ be if we're interested in the wealth channel of the crash?

Ideally, $x_s$ could be the fraction of all wealth tied to the stock market and its fluctuations per state.
This would allow us to estimate the exposure of wealth to the stock market fluctuations, and then
study the impact of the crash on consumption through this "wealth channel".
This way, the $x_s$ would be an appropriate measure of the exogenous variation in stock market exposure.
Note that having data at the household level might be more relevant if we want to observe the consumption
of the representative american household (state level aggregate data might overestimate the size of stock
market wealth in their portfolio).



### 6. Construct $x_s$ and explain the idea behind this measure:

```{r}
dta6 <- dta4 %>%
  mutate (x = dividend_income / total_income)

summary(dta6$x)
```

The idea behind this measure is to estimate the exposure of to the stock market crash by
using the portion of income linked to the stock market. To do so, we compute the share of dividend
income in total income at the state level. Dividends correspond to the distribution of firms' profits
to stockholders; they are closely tied to asset prices, thus we expect fluctuations in the stock
market to affect them. That is why we use it as a proxy for the exposure to the stock market. 
In Chodorow-Reich et al. (2021), the authors estimate an MPC of 3.2 cents of dollar for stock wealth
annually. These findings support the idea of a correlation between the exposure to the stock market
and consumption and gives us reasons to think that measuring this exposure could help us understand
the impact of the stock market crash on consumption.


### 7. Explain the idea behind the regression. 
### What are the identification concerns?

The idea behind this regression is to study the effect of a stock market crash
on the way market exposure affected consumption, using an interaction term, while
controlling for both state and time fixed-effects.

We use car sales at the state level as a proxy for consumption $y_{st}$ (independent variable).
According to Romer (1990), stock market crash caused consumers to "delay current spendings" 
on durable goods, lowering overall consumption. Cars are good examples of this type of goods,
thus observing changes in car sales might be considered as a good way to estimate changes in
consumption behavior.
$x_s$ (independent variable) captures the exposure of state incomes to 
fluctuations in the stock market; computed by dividing the proportion of total
income from dividends in the state by the state total incomes.
The interaction term enables to examine whether the relationship between the target
parameter (consumption) and the independent variable (exposure to the stock
market) changes with the stock market crash thanks to the dummy $D_t$.
Here, the beta coefficient is the difference in slope between exposure before and 
after the crash. In terms of interpretation, if the coefficient is different
from zero and statistically significant, then we can assume the crash had an impact on the way 
exposure of to the stock market affected consumption.
A time fixed-effect is added to control for potential time trends, while the
state fixed-effect should control for potential characteristics
that are specific to states, that we do not observe.
These fixed-effects are designed to eliminate omitted variable bias by 
excluding unobserved variables that evolve over time or are specific
to some states but are constant across entities.

There might be several identification concerns regarding this regression.
First, we know that the income data come from federal tax returns at a time 
when only a small fraction of the population paid federal income tax. Thus,
our sample would not be representative of the population (the representative 
american consumer) and we would face a selection bias.
Secondly, we have to consider possible issues related to our proxy;
there could have been changes in car sales associated with neither 
the crash nor the exposure to income, creating a potential ommitted variable bias.
Thirdly, the wealth channel we observe might have a problem of endogeneity.
As mentioned in Chodorow-Reich et al. (2021), stock prices are forward looking. This
would mean that if agents anticipate a fall decline in future economic fundamentals, it
lead to a fall in stock returns and spendings. In this case, the initial decrease in 
consumption could worsen the anticipations and further decrease dividends.
Finally, measurement errors can always be a concern, there might have been 
mistakes during tax filings and fiscal evasion.


### 8. Regression (1)

```{r}
dta8 = dta6 %>% #we prepare the data for regression
  filter(year %in% c(1929, 1930)) %>% #keeping only observations from 1929 and 1930
  mutate (log_car = log(car_sales)) %>% #create a variable for the log of car sales
  mutate (time = as.Date(paste(year, as.numeric(month), "01", sep = "-"))) #

#We use two regression methods to check our results:
reg8_felm = felm(log_car ~ x*crash | time + state, data = dta8)
summary(reg8_felm)

reg8_feols = feols(log_car ~ x*crash | time + state, data = dta8)
summary(reg8_feols)
```

$\beta x_s D_t$ captures the interaction effect between the state's exposure to the stock 
market ($x_s)$, and whether or not the crash happened ($D_t$) on consumption ($y_{st}$). 
$\beta$ mesures whether the effect of state's exposure to the stock market on consumption
(measured with the log of car sales) changes with the crash.
Here we find the same coefficient across both regression such that $\beta = 1.311$. We could interprete
it as: the effect of exposure to the stock market on car consumption would have increased 
after the stock market crash. However, we get from reg8_feols that our coefficient is not statistically
significant, which prevents us from concluding that the crash had any effect on consumption through
the wealth channel described in Chodorow-Reich et al. (2021).
For comparison, in Temin (1976), the author found there was an actual effect of the crash and drop in consumers'
spendings but that it was fairly small.

Note there are no coefficients on exposure ($x_s$) and whether or not the crash happened (crash).
As indicated in summary(reg8_feols), both variables are removed because of collinearity.
We can assume they were removed by controlling the fixed-effect. Indeed, $x_s$ stays constant across
each state while $D_t$ affects all states uniformly across time.


### 9. Assuming our measure of exposure to the stock market is the right one,
### would regression (1) be appropriate to capture the wealth channel of the crash?
### Would it capture the uncertainty channel that Romer writes about?

The wealth channel refers to the effect of the crash on households' wealth, which in turn affects consumption.
Assuming our measure of exposure is the right one and if consider we car sales (a durable good) to be a good
proxy for consumption, we might think this regression is appropriate to capture the wealth channel of the crash, 
keeping in mind we operate at the state level.
However, it is still difficult to disentengle the fall  in consumption due to wealth exposure to the 
stock market and to uncertainty. Indeed, as described in Chodorow-Reich et al. (2021), the main challenge when 
studying the wealth channel is that stock prices are forward-looking.  As a result, agents might decide to cut 
spendings not because of  wealth destruction but because they anticipate a "negative stock return and 
a subsequent decline in household spending and employment." In this case, an initial fall in stock prices might
worsen expectations due to uncertainty on future wealth/income, causing agents to consume less.
Romer describes the uncertainty channel as the immediate drop in consumer spending caused by a
temporary increase in income uncertainty. According to this uncertainty hypothesis, there should be an
"inverse relationship between consumer spending on durable goods" (here cars) "and uncertainty about future
income".
Thus, this regression would not fully capture Romer's uncertainty channel since it is not direcly tied to
the loss of wealth but rather the agents' behavior in an uncertain environment. Some of it might be
reflected in the interaction term since states with greater exposure might react more in this setting,
but it is not specifically designed to measure this channel.


### 10. Potential problem of income data coming from federal tax returns:

As we mentioned earlier, having only 5% of the population paying federal income tax in this
period might create a strong selection bias, as those paying these taxes might not be representative
of the average american consumer. In Temin (1976), the wealth effect is estimated to be fairly small,
one pf the reason being that stock prices only represented a small fraction of total wealth (and was 
more likely to be detained by richer households). Here we might overestimate the fraction of wealth 
tied to the stock market.


### 11. BONUS: variant of (1)


```{r}
dta11 <- dta8 %>%
  mutate(
    # Create a month-year string for easier handling (e.g., "1929-01", "1930-12")
    month_year = paste(year, sprintf("%02d", month), sep = "-")
  ) %>%
  filter(month_year != "1929-10") %>%  # Remove October 1929
  mutate(MD = as.factor(month_year))

reg11 <- feols(log_car ~ x * MD | time + state, data = dta11)
summary(reg11)

```