---
title: "Assignment 1 - Economic History"
output:
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages and Environment

```{r, results='hide'}
library(openxlsx)
library(tidyr)
library(dplyr)
library(lfe)
library(fixest)
library(jsonlite)

rm(list=ls())
setwd("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History")

```

### 2.  Download the Excel files

```{r}
car = read.xlsx("StateNewCarRegistrations.xlsx")
stock = read.xlsx("stock_income.xlsx")

head(car)
head(stock)
```

### 3. Create a single dataset.

```{r}
car_long <- pivot_longer(car, #we start by pivoting the car dataset
                               cols = -c(month, year), 
                               names_to = "state", 
                               values_to = "car_sales")
head(car_long)



dta3 <- merge(stock, car_long, by = "state", all.x = TRUE)%>%
  filter(!is.na(year)) %>% #We merge both datasets by state, filter for NAs (and remove the "total" observations)
  group_by(state)%>% #We group observations by state
  mutate(month = as.numeric(month)) %>% #We turn the class of month from character to numeric
  arrange(year, month, .by_group = TRUE) #This allows us to arrange observations by year and month within state groups

head(dta3) #The dataset seems to corresponds to the exemple given.
```


### 4. Create a post-crash dummy

```{r}
dta4 <- dta3 %>%
  mutate (crash = ifelse(year < 1929 | (year == 1929 & month <= 10), 0, 1))
```


### 5.  In an ideal world, what should $x_s$ be if we're interested in the wealth channel of the crash?

The $x_s$ should capture the impact of the exposure to the stock market 
crash at the state level on all the wealth of the households, being an appropriate measure
of the exogenous variation in stock market exposure. But such exposure is difficult to measure,
which is also why we use state-level data here.
This variable would need to reflect the pre-crash financial exposure of each 
state in a way that captures potential heterogeneous treatment effects from 
the crash, but is also exogenous to consumption trends. There must be pre-treatment exogeneity;
 $x_s$ must reflect exposure before the crash.


### 6. Construct $x_s$ and explain the idea behind this measure:

```{r}
dta6 <- dta4 %>%
  mutate (x = dividend_income / total_income)

summary(dta6$x)
```

The idea behind this measure is to estimate the exposure of income to the
stock market crash  by computing the share of total income due to dividends.
In other words, it should reflect the share of agents' income that depends on 
the stockmarket since the dividends corresponds to the reward of holdings assets.
Thus, fluctuations in the stock market will affect dividends, which will in turn
affect income and consumption.


### 7. Explain the idea behind the regression.
    What are the identification concerns.

Reg (1): studies effect of stock market crash on consumption at state level by
using state-level car sales as a proxy.
$x_s$ (independent variable) captures the exposure of state incomes to 
fluctuations in the stock market; computed by dividing the proportion of total
income from dividends in the state by the state total incomes.
The interaction term enables to examine whether the relationship between the target
parameter (consumption) and the independent variable (exposure to the stock
market) changes with the stock market crash.
Here, beta coefficient is the difference in slope between exposure before and 
after the crash. In terms of interpretation, if the coefficient is different
from zero, then we can assume had an impact on the way exposure of income
to the stock market affected consumption.

A time fixed-effect is added to control for potential time trends, while the
state fixed-effect should control for potential characteristics
that are specific to states, that we do not observe.
These fixed-effects are designed to eliminate omitted variable bias by 
excluding unobserved variables that evolve over time or are specific
to some states but are constant across entities.

#Identification concerns: endogeneity (check Romer)
#Exposure may not be constant over the analysed period.



### 8. Regression (1)

```{r}
dta8 = dta6 %>% #we prepare the data for regression
  filter(year %in% c(1929, 1930)) %>% #keeping only observations from 1929 and 1930
  mutate (log_car = log(car_sales)) %>% #create a variable for the log of car sales
  mutate (time = as.Date(paste(year, as.numeric(month), "01", sep = "-"))) #

#We use two regression methods:
reg8_felm = felm(log_car ~ x*crash | time + state, data = dta8)
summary(reg8_felm)

reg8_feols = feols(log_car ~ x*crash | time + state, data = dta8)
summary(reg8_feols)
```

#Present and explain your results



### 9. Assuming our measure of exposure to the stock market is the right one,
    would regression (1) be appropriate to capture the wealth channel of the crash?
    Would it capture the uncertainty channel that Romer writes about?

#Reg (1) appropriate to capture wealth channel?

#capture uncertainty channel (Romer)



### 10. Potential problem of income data coming from federal tax returns:




### 11. BONUS: variant of (1)