---
title: "Economic History - Assignment 2"
author: "Lionel Chambon"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(lpirfs)
library(readxl)
library(fixest)
library(data.table)
library(modelsummary)
library(pROC)
library(haven)
library(car)
library(stargazer)
library(aod)
library(pROC)
library(countrycode)

```

## Part A

```{r}

jst = read_xlsx("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History/A2/JSTdatasetR6.xlsx")

tibble(jst)
names(jst)

```

### Model Estimation

First, we have to generate the log changes and the lags of the private credit 
supply variable.

```{r}

jst <- jst %>%
  arrange(country, year) %>%
  group_by(country) %>%  
  mutate(
    rcredit = (tloans/cpi),
    rgdp = (gdp/cpi),
    log_credit = log(rcredit),  
    d_log_credit = log_credit - lag(log_credit, 1),  
    credit_lag1 = lag(d_log_credit, 1),  
    credit_lag2 = lag(d_log_credit, 2),  
    credit_lag3 = lag(d_log_credit, 3),  
    credit_lag4 = lag(d_log_credit, 4),  
    credit_lag5 = lag(d_log_credit, 5) 
  ) %>%
  ungroup()

```

Next, estimating the logit model:

```{r}

logit_a = feglm(crisisJST ~ credit_lag1 + credit_lag2 + credit_lag3 + credit_lag4 + credit_lag5 | country,
                data = jst, 
                family = binomial(link = "logit"))


```

Note that R automatically removes NA observations of the lags.

Displaying results:

```{r}

summary(logit_a)

```

This tells us that only the second lag of private credit supply is individually
significant in this specification. This is consistent with the results from
Schularick and Taylor (2012).

Now, conducting the joint tests:

```{r}

hypothesis1 <- c("credit_lag1 = 0", "credit_lag2 = 0", "credit_lag3 = 0", "credit_lag4 = 0", "credit_lag5 = 0")

wald1 <- linearHypothesis(logit_a, hypothesis1)

print(wald1)

```

The p-value is well below $0.05$, so we can reject the null hypothesis that
lags of log changes in credit supply are jointly insignificant.

Next, we load the BVX crisis data.

```{r}

bvx <- read_dta("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History/A2/BVX replication kit/BVX replication kit/data/additional data/Narrative Crisis List, Panics List, and BVX List.dta")

```

This looks a bit messy to merge. First, I identify all crisis years by country,
before expanding the time series to match the JST dataset and creating an 
appropriate BVX dummy.

```{r}

bvx_by_country <- bvx %>%
  filter(!is.na(revised)) %>%
  group_by(country) %>%
  summarize(Crisis_Years = list(revised))

print(bvx_by_country)

year_range <- 1870:2020

bvx_ts <- bvx_by_country %>%
  select(country) %>%
  distinct() %>%
  expand(country, year = year_range) %>%
  left_join(bvx_by_country, by = "country") %>%
  rowwise() %>%
  mutate(BVX = ifelse(year %in% unlist(Crisis_Years), 1, 0)) %>%
  ungroup() %>%
  select(country, year, BVX)

print(bvx_ts)

```

I check by hand for Argentina and Australia that the dates match.
Now merging:

```{r}

jst <- jst %>%
  left_join(bvx_ts, by = c("country", "year"))

print(jst)

```

Based on a manual-double check of crisis dates and dummies, this looks good.

Re-estimating the model:

```{r}

logit_b = feglm(BVX ~ credit_lag1 + credit_lag2 + credit_lag3 + credit_lag4 + credit_lag5 | country,
                data = jst, 
                family = binomial(link = "logit"))

summary(logit_b)

```

In this model, the second *and* fifth lag are highly significant.

```{r}

hypothesis2 <- c("credit_lag1 = 0", "credit_lag2 = 0", "credit_lag3 = 0", "credit_lag4 = 0", "credit_lag5 = 0")

wald2 <- linearHypothesis(logit_b, hypothesis2)
print(wald2)

```

Similarly to before, we can reject the null that credit lags are jointly
insignificant.

Next, we repeat the exercise using the five-year change in the credit-to-GDP
ratio as a predictor.

First, we have to generate the credit-to-GDP ratio and the five-year change:

```{r}

jst <- jst %>%
  arrange(country, year) %>%
  group_by(country) %>%  
  mutate(
    credit_to_gdp = (rcredit/rgdp),  
    five_year_change = credit_to_gdp - shift(credit_to_gdp, 5, type = "lag")
  ) %>%
  ungroup()

```

Re-estimating the model: 

```{r}

logit_a2 = feglm(crisisJST ~  five_year_change | country,
                data = jst, 
                family = binomial(link = "logit"))


```

Displaying results:


```{r}

summary(logit_a2)

```


The coefficient for the five-year change in the credit-to-GDP ratio is
positive and highly significant.

### Model Evaluation

First, we split the sample along observations before and after 1984.

```{r}

jst_pre1984 <- jst %>% filter(year <= 1984)
jst_post1984 <- jst %>% filter(year > 1984)

in_sample_model = feglm(crisisJST ~ credit_lag1 + credit_lag2 + credit_lag3 + credit_lag4 + credit_lag5 | country,
                data = jst_pre1984, 
                family = binomial(link = "logit"))

summary(in_sample_model)

```

The in-sample estimation produces a highly significant and positive coefficient.

```{r}

outof_sample_model = feglm(crisisJST ~ credit_lag1 + credit_lag2 + credit_lag3 + credit_lag4 + credit_lag5  | country,
                data = jst_post1984, 
                family = binomial(link = "logit"))

summary(outof_sample_model)

```

The coefficient is even more significant for the out-of-sample model.

Now, we use the in-sample estimation for the post-1984 prediction.

```{r}

jst_pre1984$predict_in_sample <- predict(in_sample_model, 
                                             newdata = jst_pre1984, 
                                             type = "response")

jst_post1984$predict_outof_sample <- predict(in_sample_model, 
                                             newdata = jst_post1984, 
                                             type = "response")

```

Next, we compute ROC curves:

```{r}

roc_in <- roc(jst_pre1984$crisisJST, jst_pre1984$predict_in_sample)
auc(roc_in)

roc_out <- roc(jst_post1984$crisisJST, 
                         jst_post1984$predict_outof_sample)
auc(roc_out)

roc_data_in <- data.frame(
  FPR = 1 - roc_in$specificities,
  TPR = roc_in$sensitivities,
  Thresholds = roc_in$thresholds,
  Model = "In-sample (AUC: 0.71)"
)

roc_data_out <- data.frame(
  FPR = 1 - roc_out$specificities,
  TPR = roc_out$sensitivities,
  Thresholds = roc_out$thresholds,
  Model = "Out-of-sample (AUC: 0.64)"
)

roc_data <- rbind(roc_data_in, roc_data_out)

roc_plot <- ggplot(roc_data, aes(x = FPR, y = TPR, color = Model)) +
  geom_line(linewidth = 1.2) +  
  labs(title = "ROC Curves: In-sample (pre-1984) vs Out-of-sample (post-1984) models", 
       subtitle = "Predictor: 5-year change in credit-to-GDP ratio",
       x = "False Positive Rate (FPR)", 
       y = "True Positive Rate (TPR)", 
       color = "Model") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("blue", "red")) +  
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray")  

print(roc_plot)


```

Next, we want to repeat the estimation using a) money and b) public debt
as predictors.


```{r}

jst <- jst %>%
  arrange(country, year) %>%
  group_by(country) %>%  
  mutate(
    rmoney = (money/cpi),
    log_money = log(rmoney),  
    d_log_money = log_money - lag(log_money, 1),  
    money_lag1 = lag(d_log_money, 1),  
    money_lag2 = lag(d_log_money, 2),  
    money_lag3 = lag(d_log_money, 3),  
    money_lag4 = lag(d_log_money, 4),  
    money_lag5 = lag(d_log_money, 5) 
  ) %>%
  ungroup()

logit_money = feglm(crisisJST ~ money_lag1 + money_lag2 + money_lag3 + money_lag4 + money_lag5 | country,
                data = jst, 
                family = binomial(link = "logit"))

summary(logit_money)

```

The coefficient on lag2 is positive and is significant at the 10%-level.

```{r}
jst <- jst %>%
  arrange(country, year) %>%
  group_by(country) %>%  
  mutate(
    d_debt = debtgdp - lag(debtgdp, 1),  
    debt_lag1 = lag(d_debt, 1),  
    debt_lag2 = lag(d_debt, 2),  
    debt_lag3 = lag(d_debt, 3),  
    debt_lag4 = lag(d_debt, 4),  
    debt_lag5 = lag(d_debt, 5) 
  ) %>%
  ungroup()

logit_debt = feglm(crisisJST ~ debt_lag1 + debt_lag2 + debt_lag3 + debt_lag4 + debt_lag5 | country,
                data = jst, 
                family = binomial(link = "logit"))

summary(logit_debt)

```

The coefficient of the debt-to-GDP ratio on lag1 is negative and significant at the 10-% confidence level.
The coefficient of the debt-to-GDP ratio on lag4 is negative and significant at the 1-% significance level.

Now, plotting the ROC curves. First, we run the predictions for both
predictors and create the dataframe required for plotting:

```{r}

jst$predict_money <- predict(logit_money, 
                             newdata = jst, 
                             type = "response")

jst$predict_debt <- predict(logit_debt, 
                            newdata = jst, 
                            type = "response")

roc_money <- roc(jst$crisisJST, jst$predict_money)
auc_money = round(auc(roc_money), 2)

roc_debt <- roc(jst$crisisJST, jst$predict_debt)
auc_debt = round(auc(roc_debt), 2)

roc_moneyvdebt <- rbind(
  data.frame(
    FPR = 1 - roc_money$specificities,
    TPR = roc_money$sensitivities,
    Model = paste("Money Model (AUC:", auc_money, ")")
  ),
  data.frame(
    FPR = 1 - roc_debt$specificities,
    TPR = roc_debt$sensitivities,
    Model = paste("Debt Model (AUC:", auc_debt, ")")
  )
)

auc_money
auc_debt


```

Finally, we obtain:

```{r}

moneyvdebt_plot <- ggplot(roc_moneyvdebt, aes(x = FPR, y = TPR, color = Model)) +
  geom_line(size = 1.2) +
  labs(
    title = "ROC Curves for Money and Public Debt as Predictors",
    x = "False Positive Rate (FPR)",
    y = "True Positive Rate (TPR)"
  ) +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  facet_wrap(~ Model) +  
  theme(legend.position = "none") 

print(moneyvdebt_plot)

```

Let's recall the previous model for comparison:

```{r}

print(roc_plot)

```



## Part B

First, we reload the required datasets, clean and merge them.

```{r}

df1 = read_xlsx("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History/A2/JSTdatasetR6.xlsx")

df2 = read_dta("C:/Users/etien/OneDrive/Documents/GitHub/Econ-History/A2/RecessionDummies.dta")

# Ensure consistent types for joining columns
df1 <- df1 %>%
  mutate(iso = as.character(iso), year = as.integer(year))

df2 <- df2 %>%
  mutate(iso = as.character(iso), year = as.integer(year))

# Replace NA values in numeric columns of df2 with 0
df2 <- df2 %>%
  mutate(across(where(is.numeric), ~replace_na(., 0)))

# Perform the left join
merged_data <- df1 %>% 
  left_join(df2, by = c("iso", "year"))

# Inspect the merged data
head(merged_data)

```
Create outcome variable

````{r}
data_set <- merged_data %>% 
  mutate(lrgdppc = log(rgdpbarro)*100)

lp_reg <- data_set %>%
  select(iso, year,lrgdppc , N, `F`) %>%
  mutate(
    N = ifelse(is.na(N), 0, N),
    `F` = ifelse(is.na(`F`), 0, `F`)
  )

````

IRF for normal recession
````{r}

normal <- lp_lin_panel(
  data_set = lp_reg,
  endog_data = "lrgdppc",        # Dependent variable
  cumul_mult = TRUE,         # Cumulative multipliers
  shock = "N",               # Normal recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",  # Fixed effects model
  c_fd_exog_data = "F", #Control for F
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Plot impulse response functions (IRFs)
plot(normal)

````



````{r}
financial <- lp_lin_panel(
  data_set = lp_reg,
  endog_data = "lrgdppc",    # Dependent variable
  cumul_mult = TRUE,         # Cumulative multipliers
  shock = "F",               # Normal recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",    # Fixed effects model
  c_fd_exog_data = "N", #Control for N
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Plot impulse response functions (IRFs)
plot(financial)
````















